---
title: "R Notebook - Ramping Up"
output:
  html_document:
    df_print: paged
---

# 'ls()'

CNTRL-SHIFT-ENTER
```{r}
ls();
ls(all.names=TRUE);
```

# '.Machine'
```{r}
print("== .Machine ==");
str(.Machine);
```

# '.Platform'
```{r}

print("== Platform ==");
str(.Platform);
```


# '.GlobalEnv'
```{r}

print("== GlobalEnv ==");
str(.GlobalEnv);
```


# 'R.Version()'
```{r}

print("== VERSION ==");
R.Version();
```

# 'Sys.getenv()'
```{r}

print("== SYSTEM ENVIRONMENT ==");
Sys.getenv();
```

# '.Options'
```{r}

# str(.__C__aov);
print("== OPTIONS ==");
str(.Options);

```

# '.Last.value'
```{r}
foo = function(x) { 3*x^2 - 2*x; }
.Last.value;
print("== FORMALS ==");
formals(foo);
print("== BODY ==");
body(foo);
.Last.value;
```

# Manual change '.Machine'
```{r}
ls();
ls(all.names=TRUE);

.Machine$monte = "says-hi";
.Platform$monte = "says=hi";

ls();
ls(all.names=TRUE);
```


# 'set.seed(NULL)'
```{r}
set.seed(NULL);
.Last.value;
(ssNULL = runif(100, 1, 10));
```

```{r}
ls(all.names=TRUE);
```

## '.Random.seed'
```{r}
str(.Random.seed);
```
# 'set.seed(123)'
```{r}
set.seed(123);
.Last.value;
(ss123 = runif(100, 1, 10));
```

## '.Random.seed'
```{r}
str(.Random.seed);
```

# Seed 'loss'
```{r}
set.seed(123);
x = y = runif(10, 1, 5);
x == y;
identical(x,y);
x;
```

## 'while' loop
```{r}
delay = 0;
while( identical(x,y) )
  {
  delay = 1 + delay;
  Sys.sleep(1);
  if(delay %% 10 == 0) { print(paste0("Delay: ",delay)); }
  y = runif(10, 1, 5);
  }

delay;
y;
```

## Seed reset
```{r}
set.seed(123);
y = runif(10, 1, 5);
y;
```

## 'for' loop
```{r}
set.seed(123);
for(i in 1:10)
{
  y = runif(10, 1, 5);
  print(y);
}
```

## CORRECT syntax
```{r}
for(i in 1:5)
{
  set.seed(123); 
  y = runif(10, 1, 5);
  print(y);
}
```


# 'get.seed'
The purpose of seeding is for replicability across devices and time.  If you can save the seed when you set.seed(store_seed), then you could get.seed

* does an inverse exist
* does it meet the primary "uses cases"
      - set/get seed
      - multiple memories
* is it 'intuitive' with options for different 'use cases' ... SEE 'rand()'

## '.humanVerse'
The humanVerse will track the memory of seeds, and other things.  An initialization routine is part of the library, so this will auto-load.  For now,

```{r}
.GlobalEnv$.humanVerse = list();


#' initSeedMemory
#'
#' @param purge.memory If TRUE, this memory is erased and reset.
#'
#' @return NULL (nothing)
#' @export
#'
#' @examples
#' # initSeedMemory();
#' # setSeed(); getSeed(); initSeedMemory(purge.memory = TRUE); getSeed();
initSeedMemory = function(purge.memory = FALSE, verbose = FALSE)
  {
  if(!exists(".humanVerse")) { initMemory(); }

  if(!exists("seed", .GlobalEnv$.humanVerse) || purge.memory)
    {
    if(verbose)
      {
	    cat("humanVerse::initSeedMemory ... initializing list '.humanVerse[[\"seed\"]]'", "\n");
      }
    .GlobalEnv$.humanVerse[["seed"]] = list();
    }
  }

```

## 'initSeedMemory' in '.humanVerse'
```{r}
initSeedMemory( verbose = TRUE);
```

## '.humanVerse' is hidden
```{r}
str(.humanVerse);
```



## functions depend on 'rand'
```{r}

rand = function(min = -1*as.integer(Sys.time()), max = as.integer(Sys.time()), n = 1, method = "high-low", sample.replace = TRUE, seed = NULL, attributes=NULL)
    {
	# if(is.null(seed)) { setSeed(NULL, "rand"); my.seed = getSeed("rand"); } else { my.seed = seed; }
    me = substr( trimMe( tolower(method) ), 1, 2);
	n = as.integer ( n );
	if(is.na(n) || n < 1)
		{
		warning( paste0('Bad value for n "', n, '" in function [rand]', "\n", "Setting n=1") );
		n = 1;
		}
	res = NULL;
    if(me == "fl")  # floor method
      {
	  if(!is.null(seed)) { set.seed(seed); }
      res = ( as.integer( floor( stats::runif(n, min = min, max = (max + 1) ) ) ) );
      }
    else if(me == "sa")  # sample method
      {
      if(!sample.replace)
        {
        len = (max - min) + 1;
        if(len < n)
          {
          warning( "sample.replace forced to TRUE" );
          sample.replace = TRUE;
          }
        }
	  if(!is.null(seed)) { set.seed(seed); }
      res = ( sample(min:max, n, replace = sample.replace) );
      }
	else
		{
		if(me != "hi")  # high-low method
			{
			warning( paste0('Bad value for method "', method, '" in function [rand]', "\n", "Setting method='high-low'") );
			}
		# DEFAULT will run 
		if(!is.null(seed)) { set.seed(seed); }
		res = ( as.integer(( (max + 1) - min) * stats::runif(n) + min) );
		}
		
	if(!is.null(attributes))
		{
		res = setAttribute('min', min, res);
		res = setAttribute('max', max, res);
		res = setAttribute('seed', seed, res);
		res = setAttribute('method', method, res);	
		}
	
	res;
    }




```

## HELPER functions
```{r}

trimMe = function(str, side="both", method="stringi")
  {
  # sides = c("both", "left", "right")
  # methods = c("cpp", "stringi", "base")
  side = tolower(side);
  ###m = substr(tolower(method),1,1);  # is this slowing me down?
  # stringr::str_trim(str);
  # if(!is.element(side,sides)) { stop("option for 'side' must be one of:  both, left, right"); }
  # set default to both

  # dump stringr and go to stringi
  # dump tidyverse altogether
  # review httr and build appropriate functions
  # use base whenever possible, then the best packages whenever possible
  # define best by "least associated with new grammar tactics"
  # new grammar is arbitrary and not c-based
  # human-readable functions with variadic inputs is not new grammar
# lsf.str("package:stringi")
# ls("package:stringi")

	if(method == "cpp" && exists("cpp_trim"))  # humanVerse ... HVcpp
			{
			res = switch(side,
							  "left"  = cpp_ltrim(str),
							  "right" = cpp_rtrim(str),
							  "both"  = cpp_trim(str),
					cpp_trim(str)
					);
			return (res);
			}
		
		
  if( method == "stringi" && isTRUE(requireNamespace("stringi", quietly = TRUE)) )
    {
    res = switch(side,
						  "left"  = stringi::stri_trim_left(str),
						  "right" = stringi::stri_trim_right(str),
						  "both"  = stringi::stri_trim_both(str),
				stringi::stri_trim_both(str)
				);
	return (res);
	}
	
	
		
    res = switch(side,
						  "left"  = gsub("^\\s+", "", str),
						  "right" = gsub("\\s+$", "", str),
						  "both"  = gsub("^\\s+|\\s+$", "", str),
                  gsub("^\\s+|\\s+$", "", str)
                  );
    return (res);
  }



setAttribute = function(myAttribute, myValue, myObj)
	{
	attributes(myObj)[[myAttribute]] = myValue;
	myObj;  # no object referencing, so I must return
	}


```

## 'initSeed'
```{r}

#' initSeed
#'
#' @param ... Parameters that can be passed onto the 'rand' function
#'
#' @return A single integer
#' @export
#'
#' @examples
#' # initSeed();
#' # initSeed(1, 10^5, method="floor");
initSeed = function(...)
  {
  rand(...)[1];  # if they wrongly passed in n > 1, we only return [1]
  }

```

```{r}
initSeed();
```

## 'setSeed'
```{r}

#' setSeed
#'
#' @param seed
#' @param key
#' @param ...
#' @param args.set
#' @param print.seed
#' @param verbose
#'
#' @return
#' @export
setSeed = function(seed, key, ..., args.set = list(), print.seed = TRUE, verbose = FALSE )
  {
  # can I have two ellipses in a single function?
    # one for initSeed
    # one for set.seed
    # nope, argslist
  if( missing(key) ) { key = "last"; }
  if(is.null(seed))
    {
    seed = initSeed(...);
    if(verbose)
      {
      cat("setSeed :: generating new integer seed ... ", "\t", seed, "\n");
      }
    initSeedMemory();
    }
  
  if( !exists("kind", args.set) )         { kind = NULL; }
  if( !exists("normal.kind", args.set) )  { normal.kind = NULL; }
  if( !exists("sample.kind", args.set) )  { sample.kind = NULL; }
  
  my.seed = seed;
	my.seed = setAttribute("kind", kind, my.seed);
	my.seed = setAttribute("normal.kind", normal.kind, my.seed);
	my.seed = setAttribute("sample.kind", sample.kind, my.seed);
	
  .GlobalEnv$.humanVerse[["seed"]][[key]] = my.seed;
  if(verbose)
      {
      cat("setSeed :: global value stored [key] = ",key," ... [seed] = ",seed, "\n");
      }
	  
  if(verbose)
      {
      cat("setSeed :: calling base::set.seed with seed ... ", seed, "\n");
      }

  if(print.seed)
	{
	cat("\n setSeed: ", seed, "\n");
	}
  
# call set.seed at very end ...
  set.seed(seed, kind=kind, normal.kind=normal.kind, sample.kind=sample.kind);
# we don't return the seed, but we could?  getSeed will handle this ... 
  # invisible(seed);
  }


```

### 'setSeed(NULL)'
```{r}
setSeed(NULL);  # behaves like the 'normal set.seed() function'
sSNULL = runif(100, 1, 10);
```

#### EQUAL?
```{r}
identical(ssNULL, sSNULL);
```
They have different seeds since 'NULL' doesn't something blind under the hood.

### 'setSeed(123)'
```{r}
setSeed(123);
(sS123 = runif(100, 1, 10));
```

#### EQUAL?
```{r}
identical(ss123, sS123);
```
This is the whole reason we have the idea of 'set.seed' so we can get REPLICABLE results across time, space, and computing machinery.

Where is the inverse of setSeed; e.g., getSeed

## 'getSeed'
```{r}

#' getSeed
#'
#' @param key
#' @param verbose
#'
#' @return
#' @export
#'
#' @examples
getSeed = function(key, keep.attributes = FALSE, verbose = FALSE)
  {
  # I could create a "keyed" list of memory, not just last ...
  if( missing(key) ) { key = "last"; }
  if(verbose) { cat("getSeed :: looking up key ... ", "\t", key); }
  if(exists(key, .GlobalEnv$.humanVerse[["seed"]]))
    {
    if(verbose) { cat("\n\t ... found with value: ", "\t", .GlobalEnv$.humanVerse[["seed"]][[key]], "\n"); }
    my.seed = .GlobalEnv$.humanVerse[["seed"]][[key]];
	if(!keep.attributes) { my.seed = as.integer(my.seed); }
	my.seed;
    } else { FALSE; }
  }



```

### getSeed() ... last value
```{r}
getSeed();
```

## Good algebra
Now we can set/get a seed ... QED

```{r}
setSeed(NULL);
sS1 = rnorm(100);
getSeed();
setSeed(getSeed());
sS2 = rnorm(100);
```

### EQUAL?
```{r}
identical(sS1,sS1);
```

## Multiple memories
We have a key notation to remember as many as we want.  If we don't pass a key, we only have the 'last one' as a memory.

```{r}
setSeed(NULL, "uniform");  # this is how 'random' research should be done
setSeed(NULL, "normal");   # pass in a random value


(u.seed = getSeed("uniform"));
(n.seed = getSeed("normal"));


```

### SIMULATE ME
```{r}
sim = NULL;

I = 1:5;
J = 1:6;
N = 100;
U = c(-2,2);

for(i in I)
{
setSeed(NULL, "uniform");
u.dist = runif(N, U[1], U[2]);
(u.seed = getSeed("uniform"));
  for(j in J)
  {
  setSeed(NULL, "normal");
  n.dist = rnorm(N);
  (n.seed = getSeed("normal"));
  
  u.median = median(u.dist);
  n.median = median(n.dist);
  row = c(i, j, u.seed, u.median, n.seed, n.median, median(u.dist*n.dist), u.median*n.median );
  sim = rbind(sim,row);
  }
}

sim = data.frame(sim);
colnames(sim) = c("i", "j", "u.seed", "u.median", "n.seed", "n.median", "median(u*n)", "median(u)*median(n)");

sim = setAttribute("details", list("I"=I,"J"=J,"N"=N,"U"=U), sim);

sim;
```

We have stored the 'seeds' so we could rebuild this entire simulation from the data.  

**GOOD DATA SCIENCE**

```{r}
str(sim);
```

I set truly "PSEUDO-RANDOM" seeds, not just using the classic 'set.seed(123);'

If I had the code base AND the data, I could rebuild the entire simulation.


This is what **TRUE OPEN DATA SCIENCE** should look like.


```{r}
getAllAttributes = function(myObj)
	{
	attributes(myObj);
	}
```

### REPLICATION CHECK
```{r}
sim.check = NULL;

details = getAllAttributes(sim)$details;

lookupSeed.I = function(n, df)
  {
  r = which( (as.numeric(df["i"] == n)) == 1)[1];
  df[r,]$u.seed;
  }


lookupSeed.J = function(n, u.seed, df)
  {
  i = which( (as.numeric(df["u.seed"] == u.seed)) == 1);
  j = which( (as.numeric(df["j"] == n)) == 1);
  r = intersect(i,j)[1];
  df[r,]$n.seed;
  }

I = as.numeric( unlist( details["I"] ) );       # I = as.integer( unlist( details["I"] ) );
J = as.numeric( unlist( details["J"] ) );       # J = as.integer( unlist( details["J"] ) );
N = as.numeric( unlist( details["N"] ) );       
U = as.numeric( unlist( details["U"] ) );       

for(i in I)
{
setSeed( lookupSeed.I(i, sim), "uniform");
u.dist = runif(N, U[1], U[2]);
(u.seed = getSeed("uniform"));
  for(j in J)
  {
  setSeed( lookupSeed.J(j, u.seed, sim), "normal");
  n.dist = rnorm(N);
  (n.seed = getSeed("normal"));
  
  u.median = median(u.dist);
  n.median = median(n.dist);
  row = c(i, j, u.seed, u.median, n.seed, n.median, median(u.dist*n.dist), u.median*n.median );
  sim.check = rbind(sim.check,row);
  }
}

sim.check = data.frame(sim.check);
colnames(sim.check) = c("i", "j", "u.seed", "u.median", "n.seed", "n.median", "median(u*n)", "median(u)*median(n)");

sim.check = setAttribute("details", list("I"=I,"J"=J,"N"=N,"U"=U), sim.check);

sim.check;

```

```{r}
str(sim.check);
```

#### EQUAL?
```{r}
identical(sim, sim.check);
```

EASTER +10:  WHY ARE THEY NOT IDENTICAL?

#### EQUAL by pieces?
```{r}
identical( sim$i,       sim.check$i );
identical( sim$u.seed,  sim.check$u.seed );
identical( sim$j,       sim.check$j );
identical( sim$n.seed,  sim.check$n.seed );
```

#### EQUAL loop

```{r}
sim.names = colnames(sim);
for(sim.name in sim.names)
  {
  print(sim.name);
  # sim$i == unlist( sim["i"] ) ... almost equivalent, one is variadic
  print( identical(sim[sim.name], sim.check[sim.name]) );
  }

```
Above shows it is not an issuing with floating point.

#### EQUAL attributes?
```{r}
( attr.sim = getAllAttributes(sim) );
( attr.sim.check = getAllAttributes(sim.check) );

identical(attr.sim, attr.sim.check);

```

```{r}

names(attr.sim) == names(attr.sim.check);

```
##### Loop over names
```{r}
attr.names = names(attr.sim);

for(attr.name in attr.names)
  {
  print(attr.name);
  print( identical(attr.sim[attr.name], attr.sim.check[attr.name]) );
  }
```

##### 'details' attribute is issue
```{r}
str(attr.sim$details);
str(attr.sim.check$details);

identical(attr.sim$details, attr.sim.check$details);
```

##### Type (casting) issue
```{r}
str(attr.sim$details$I);
str(attr.sim.check$details$I);

```

There it is.  I run R which is "loosely" typed.  One time, R decides I is an integer, the other time it decides I is a number.  

```{r}
# solution ... change value on left to value on the right
I = as.numeric( unlist( details["I"] ) );       I = as.integer( unlist( details["I"] ) );
J = as.numeric( unlist( details["J"] ) );       J = as.integer( unlist( details["J"] ) );

(I = 1:5);        str(I);
(J = 1:6);        str(J);
(N = 100);        str(N);
(U = c(-2,2));    str(U);
```

```{r}
N2 = 100:100;             str(N2);
U2 = seq(-2,2, by=4);     str(U2);
U3 = seq(2,12, by=4);     str(U3);
```

I expected the sequences to be integers, but "R" doesn't have negative integers so it casts it as a "numeric"?  I could recast using as.integer on a sequence to use as part of a 'for' loop, but it appears these numerics work fine (loosely typed) in the 'for' loop but not the 'identical' check.

Under the hood it appears there are subtle differences.  Don't worry, they may change SILENTLY under the radar by the core R-dev team.  You will never know unless you carefully read the CHANGELOG.  And of course, they will tell you to RTFM!

```{r}
identical(N, N2);
```

```{r}
str(3* 100:100);
str(3.14 * 100:100);
```


A good algebra works as expected.  In the same breath you can't say "It is loosely typed" and then say "you have to be careful with your numeric typing".  

As a result, you will have such issues as demonstrated above.