---
title: 'R Notebook example: Hello World'
name: PUT --YOUR NAME-- HERE
email: PUT --YOUR EMAIL-- HERE
output:
  pdf_document: default
  html_document: default
---

# What is a library?

## Number of Cores
```{r}
# Cntrl-Alt-I is your friend

library(parallel);          # install.packages("parallel", dependencies=TRUE);
parallel::detectCores();

# Cntrl-Shift-C is also your friend

```
Can you find another library that counts cores?  Is it a core or a thread?  My laptop is a QUAD-core.

## Prime Numbers

```{r}
library(pracma);
pracma::primes(22);
```


```{r}
prime.pracmaPrimes = function(n, first=FALSE)
	{
	# this duplicates the primary logic of pracma::primes
	# by computing 'sqrt' one time, it speeds up things 'slightly'
	# it allows for firstN or fromN with first=FLAG
	gn = n;
	if(first) { gn = ceiling( n * log(n) + n * log(log(n)) ); }
	gn.sqrt = floor( sqrt(gn) );  # needs to round down so the "seq by k" doesn't break ...

	p = seq(1, gn, by = 2); # odd numbers
    q = length(p);
    p[1] = 2; 	# replace 1 with 2 (prime)
				# 9 is the first non-prime?
	if(gn >= 9)
		{
		for (k in seq(3, gn.sqrt, by = 2) )
			{
			k.idx = (k + 1)/2;
			if (p[k.idx] != 0)
				{
				# using a squared rule on indexes ?
				k2.idx = (k * k + 1)/2;

				# cat("\n", " k = ", k, "... (k+1/2) = ", k.idx, "... (k * k + 1)/2 = ", k2.idx, "\n");

				p[ seq(k2.idx, q, by = k) ] = 0;
				}
			}
		}

	p = p[p > 0];

	if(first)
		{
		p[1:n];
		} else {
				p[p < n];
				}
	}
```


```{r}
prime.pracmaPrimes(22);
```

```{r}
prime.pracmaPrimes(22, first=TRUE);
```

### Benchmarking
```{r}
library(microbenchmark);
microbenchmark::microbenchmark(
                                pracma::primes(22) , 
                                prime.pracmaPrimes(22), 
                                prime.pracmaPrimes(22, first=TRUE) 
                                );
```
One line of code is inefficient in 'pracma', can you find it?

```{r}
pracma::primes
```
The 'sqrt(n)' is not changing but is being computed several times in the 'k' sequence.

# Does 'hello world' work?

```{r}
hello = function()
  {
  print("Hello, world!")
  }
hello();
```

```{r}
source("https://raw.githubusercontent.com/MonteShaffer/MasterClassDataAnalytics/main/functions/functions-hello.R");
hello();
```

# Does Python Work?

```{r}
library(reticulate);        # install.packages("reticulate", dependencies=TRUE);
reticulate::use_python("C:/Python/Python39")
```

Pay attention to detail.  This will only work if all the building blocks are in place.
```{python}
print("hello world");
```

```{python}
def pownum(base, pow):
  return base ** pow

print(pownum(9,5));
```

```{python}
print(pownum(3,5));
```


# Does C++ Work?

```{r}
library(Rcpp);        # install.packages("Rcpp", dependencies=TRUE);
# https://stackoverflow.com/questions/64839024/
```


## Bits and Such
```{r}
Rcpp::cppFunction("long long RShift(long long a, int b) { return a >> b;}")
```

### Shifty shifting

The "right shift" operator in R is based on S+ and has some limitations when it comes to signed (negative) integers.

```{r}
y = 1732584194;
RShift(y, 16);
bitwShiftR(y, 16);
```


```{r}
y = -1732584194;
RShift(y, 16);
bitwShiftR(y, 16);
```

Libraries 'bit' and 'bit64' may be of some benefit, but beforewarned when working with bits using R.

```{r}
bitShiftR = function(x, bits, unsigned=FALSE)
  {
  if(!is.negative(x) | unsigned) { return( bitwShiftR(x,bits) ); }
  -bitwShiftR(-x,bits) - 1; #  - 1;                  # >>>
  }

is.negative = function(x, ..., tol = sqrt(.Machine$double.eps), part="Re")
  {
  more = unlist(list(...)); x = c(x, more);
  x = if(part == "Im") { x = Im(x); } else { x = Re(x); }
  x < ( -1 * tol );
  }

y = -1732584194;
RShift(y, 16);
bitwShiftR(y, 16);
bitShiftR(y, 16);

```
### Equality
```{r}
step = 0.75;
step;
```
```{r}
options(digits=22);
step == 0.75;
```

```{r}
pi;
sin(pi);
sin(pi) == 0;
# isClose(a,b);
```



### Convert decimal number to a binary string (and vice versa)

#### decbin function
```{r}
# https://stackoverflow.com/questions/6614283/converting-decimal-to-binary-in-r

dec2bin = decbin = function(decnum) 
	{
	bvect = rep(0, 1 + floor(log(decnum, 2))); # pre-populate with zeroes
	while (decnum >= 2) 
		{
		power = floor(log(decnum, 2));
		bvect[1 + power] = 1;
		decnum = decnum - 2^power;
		} 
	bvect[1] = decnum %% 2;
	paste(rev(bvect), collapse = ""); # convert to a string, reversed
	}
```


#### decbin example
```{r}
decbin(57);
```
If you write a function, you should also have its inverse.

#### bindec function
```{r}
# two names for the function
  # bin2dec is matlab
  # bindec is PHP 
bin2dec = bindec = function(binstr)
  {
  n = strlen(binstr);
	res = 0; power = 0;
	for(i in n:1) # we reversed it in the for loop
		{
		bit = as.integer(charAt(binstr,i));
		add = 0;
		if(bit == 1) { add = 2^power; }
		
		res = res + add;
		power = 1 + power;
		}
	res;
  }

```

```{r}
## bin2dec('111001');  # you may want to comment this out when you Knit-HTML as it may throw an "intentional" error.
```

#### helper functions
```{r}
strlen = function(str)
  {
  # history :: # https://en.cppreference.com/w/c/string/byte/strlen
  # http://www.cplusplus.com/reference/cstring/
  # https://en.wikipedia.org/wiki/C99
  # https://www.programiz.com/c-programming/library-function/string.h/strlen
  # vectorized ... already
  nchar( as.character(str), type="chars");
  }

charAt = function(str,idx)
  {
  substr(str,idx,idx);
  }

```

#### bindec example

```{r}
bin2dec('111001');
```

```{r}
bindec('111001');
```



#### INVERSE check
```{r}
bindec( decbin(57) );

```

```{r}
decbin( bindec('111001') );
```
```{r}
typeof( 57 );
```


```{r}
typeof( decbin(57) );
```

#### String PAD zeroes
You could left-side 'strPadLeft' with zeroes if you wanted it to be a certain bit length

```{r}
strPadLeft = function(str, final.str.len, padding="0", method="stringi")
  {
  if( isTRUE(requireNamespace("stringi", quietly = TRUE)) && method=="stringi" )
    {
    stringi::stri_pad_left(str, final.str.len, pad = padding);
    } else {
            n = strlen(str);
            r = final.str.len - n;
            if(r < 0) { stop("strPadLeft is too short!"); }
            paste0(paste(rep(padding,r),collapse=""),str);

            }
  }
```



```{r}
strPadLeft( decbin(57), 8);
```


```{r}
strPadLeft( decbin(57), 8, method="base");
```

```{r}
strPadLeft( decbin(57), 8, method="Adljblkjadlk");
```

#### Benchmarking speed

```{r}
library(microbenchmark);
microbenchmark::microbenchmark(
                                strPadLeft( decbin(57), 8) , 
                                strPadLeft( decbin(57), 8, method="base"), 
                                strPadLeft( decbin(57), 8, method="Adljblkjadlk")
                                );
```

We are also benchmarking the 'decbin' function which likely can also be improved upon.  It is a good idea to isolate what you are actually timing, but testing in context is not a bad idea.

Since the library 'stringi' is written in C++, it has some native efficiencies over the R interpreted 'base' solution. [https://cran.r-project.org/web/packages/stringi/index.html]

## Matrices with External C++ file

This will source and compile the code.  Maybe give it a minute.
```{r}
Rcpp::sourceCpp("multiply.cpp");
```

```{r}
A = matrix(rnorm(10000), 100, 100); # fully populated, 100 x 100, relatively small
B = matrix(rnorm(10000), 100, 100);

# base::matrix
```

```{r}
library(microbenchmark);
microbenchmark::microbenchmark(
                                eigenMatTrans(A),
                                A%*%B, 
                                armaMatMult(A, B), 
                                eigenMatMult(A, B), 
                                eigenMapMatMult(A, B)
                                );

```
It appears 'eigen' performs well for me; 'arma' is about equivalent to the built-in R multiplication.

Can we use sparse matrices and pass them into C++?  Can we pass an 'R' sparse matrix into a C++ function call for speed purposes?  

## Imagery

Two powerful C/C++ tools now accessible within R.
```{r}
library(magick); #install.packages("magick", dependencies=TRUE);
# https://cran.r-project.org/web/packages/magick/vignettes/intro.html#The_grid_package
# https://www.datanovia.com/en/blog/easy-image-processing-in-r-using-the-magick-package/
# https://www.imagemagick.org/discourse-server/viewtopic.php?t=18433

# 
image_content = function(x, ...)
  {
  x = magick::image_data(x, ...);
  as.integer(x);
  }

tiger         = magick::image_read_svg('http://jeroen.github.io/images/tiger.svg', width = 350);

tiger;
```

```{r}
tiger.png     = magick::image_convert(tiger, "png");
tiger.png; 
```


```{r}
tiger.matrix  = image_content(tiger.png);
dim(tiger.matrix);
# 3D array
# tiger.matrix[,,1];  # x,y, z ... z is likely RGBa = 4 layers

# tiger.matrix;
```

```{r}
library(tesseract); # install.packages("tesseract");
# https://github.com/ropensci/magick/issues/154

img.file = "iris-ocr-intro.png";
img = magick::image_read( img.file );
#img.txt = tesseract::image_ocr(img);  # renamed during LIBRARY upgrade?
img.txt = tesseract::ocr(img);

cat(img.txt);
```


```{r}
img.file = "iris-ocr.png";

img = magick::image_read( img.file );
# img.txt = tesseract::image_ocr(img);  # renamed during LIBRARY upgrade?
img.txt = tesseract::ocr(img);

cat(img.txt);
```


# Does Java Work?

Natural language processing requires java running under the hood.


```{r}
library(openNLP);  # this requires rJava ... Java
library(NLP);

sentence.a = openNLP::Maxent_Sent_Token_Annotator();
word.a     = openNLP::Maxent_Word_Token_Annotator();
```


```{r}
s = anna = "Happy families are all alike; every unhappy family is unhappy in its own way.";

sw.a       = NLP::annotate(s, list(sentence.a, word.a));

pos.a      = openNLP::Maxent_POS_Tag_Annotator(probs=TRUE);
swpos.a    = NLP::annotate(s, list(pos.a), sw.a);
```

```{r}
swpos.a.words = subset(swpos.a, type=="word");

(swpos.a.words);  
```

```{r}
# use substring to map back to the original value
substring(s, swpos.a.words$start, swpos.a.words$end);
# NOT a dataframe, so will have to further manipulate
```



```{r}
# seldom will I use an "apply" macro.
tags = sapply(swpos.a.words$features, `[[`, "POS");
sort(table(tags),decreasing = TRUE);
```




