---
title: "R Notebook - benchmark efficiencies"
output:
  html_document:
    df_print: paged
---

# Necessary Functions

CNTRL-ALT-R (run everything) 
```{r}

setAttribute = function(myAttribute, myValue, myObj)
	{
	attributes(myObj)[[myAttribute]] = myValue;
	myObj;  # no object referencing, so I must return
	}

doMedian = function(x, type=1)
  {
  xx = stats::na.omit(x);
  as.numeric(stats::quantile(xx, prob=c(0.5), type=type));
  }


strPadLeft = function(str, final.str.len, padding="0", method="stringi")
  {
  if( isTRUE(requireNamespace("stringi", quietly = TRUE)) && method=="stringi" )
    {
    stringi::stri_pad_left(str, final.str.len, pad = padding);
    } else {
            n = strlen(str);
            r = final.str.len - n;
            if(r < 0) { stop("strPadLeft is too short!"); }
            paste0(paste(rep(padding,r),collapse=""),str);

            }
  }

strlen = function(str)
  {
  # history :: # https://en.cppreference.com/w/c/string/byte/strlen
  # http://www.cplusplus.com/reference/cstring/
  # https://en.wikipedia.org/wiki/C99
  # https://www.programiz.com/c-programming/library-function/string.h/strlen
  # vectorized ... already
  nchar( as.character(str), type="chars");
  }

charAt = function(str,idx)
  {
  substr(str,idx,idx);
  }

bin2dec = bindec = function(binstr)
  {
  n = strlen(binstr);
	res = 0; power = 0;
	for(i in n:1) # we reversed it in the for loop
		{
		bit = as.integer(charAt(binstr,i));
		add = 0;
		if(bit == 1) { add = 2^power; }
		
		res = res + add;
		power = 1 + power;
		}
	res;
  }


dec2bin = decbin = function(decnum) 
	{
	bvect = rep(0, 1 + floor(log(decnum, 2))); # pre-populate with zeroes
	while (decnum >= 2) 
		{
		power = floor(log(decnum, 2));
		bvect[1 + power] = 1;
		decnum = decnum - 2^power;
		} 
	bvect[1] = decnum %% 2;
	paste(rev(bvect), collapse = ""); # convert to a string, reversed
	}
```

# Benchmark Object 'mb.res'
```{r}
library(microbenchmark);
mb.res = microbenchmark::microbenchmark(
                                        strPadLeft( decbin(57), 8) , 
                                        strPadLeft( decbin(57), 8, method="base"), 
                                        strPadLeft( decbin(57), 8, method="Adljblkjadlk"),
                                        times=100,
                                        control=list("order"="random","warmup"=2)
                                        );
mb.res;
```
# SANDBOX
```{r}
mb.res$which = as.numeric(mb.res$expr);

ids = unique(as.numeric(mb.res$expr));
names = unique(as.character(mb.res$expr));


out = NULL;

A.order = which(ids == 1); # this is the order of A
print(A.order);
A.name = names[A.order];

A = subset(mb.res, which==A.order)$time;
A.median = doMedian(A)/1000; #microseconds

row = c(1,A.order,A.name,A.median,0);
out = rbind(out, row);

n.ids = length(ids);

if(n.ids > 1)
{
  for(i in 2:n.ids)
  {
    B.order = which(ids == i); # this is the order of B
    print(B.order);
B.name = names[B.order];

B = subset(mb.res, which==B.order)$time;
B.median = doMedian(B)/1000; #microseconds

B.eff = round(100* (A.median-B.median)/A.median , 2);

row = c(i,B.order,B.name,B.median,B.eff);
out = rbind(out, row);
  }
  
  



}

out = data.frame(out);
colnames(out)=c("Index","Order","Name","Median","Delta.Efficiency");
# rownames(out)=out$Name;

out;
```

# FUNCTION
```{r}


benchmark.efficiency = function(mb.res)
	{
	mb.res$which = as.numeric(mb.res$expr);

	mb.ids = unique(as.numeric(mb.res$expr));
	mb.names = unique(as.character(mb.res$expr));

	out = NULL;

	A.order 	= which(mb.ids == 1); # this is the order of A
	A.name 		= mb.names[A.order];
	A 			= subset(mb.res, which==A.order)$time;
	A.median 	= doMedian(A)/1000; #microseconds

	row = c(1,A.order,A.name,A.median,0); # BENCHMARK, delta efficiency is 0
	out = rbind(out, row);

	n.ids = length(mb.ids);
	if(n.ids > 1)
		{
		for(i in 2:n.ids)
			{
			B.order 	= which(mb.ids == i); # this is the order of B
			B.name 		= mb.names[B.order];
			B 			= subset(mb.res, which==B.order)$time;
			B.median 	= doMedian(B)/1000; #microseconds
			B.eff 		= round(100* (A.median-B.median)/A.median , 2);

			row = c(i,B.order,B.name,B.median,B.eff);
			out = rbind(out, row);
			}
		}

	out = data.frame(out);
	colnames(out)=c("Index","Order","Name","Median","Delta.Efficiency");
	# rownames(out)=out$Name;
	rownames(out) = NULL;
		out$Median = as.numeric(out$Median);
		out$Delta.Efficiency = as.numeric(out$Delta.Efficiency);
		out$Ranking = order(out$Delta.Efficiency, decreasing=TRUE); # not to be confused with 'rank'

	out;
	}
	
	
	

```

# FUNCTION updated
```{r}


benchmark.efficiency = function(mb.res)
	{
	mb.names = levels(mb.res$expr);
	
	out = NULL;

	A.name 		= mb.names[1];
	A 			= subset(mb.res, expr==A.name)$time;
	A.median 	= doMedian(A);

	row = c(1,A.name,A.median,0); # BENCHMARK, delta efficiency is 0
	out = rbind(out, row);

	n.names = length(mb.names);
	if(n.names > 1)
		{
		for(i in 2:n.names)
			{
			B.name 		= mb.names[i];
			B 			= subset(mb.res, expr==B.name)$time;
			B.median 	= doMedian(B);
	
			B.eff 		= round(100* (A.median-B.median)/A.median , 2);

			row = c(i,B.name,B.median,B.eff);
			out = rbind(out, row);
			}
		}

	out = data.frame(out);
	colnames(out)=c("Index","Name","Median","Delta.Efficiency");
	# rownames(out)=out$Name;
	rownames(out) = NULL;
		out$Median = as.numeric(out$Median);
		out$Delta.Efficiency = as.numeric(out$Delta.Efficiency);
		#out$Ranking = order(out$Delta.Efficiency, decreasing=TRUE); # not to be confused with 'rank'
		out$Rank = rank(out$Median, ties="first");
	
	out = setAttribute("microbenchmark.unit",.Options$microbenchmark.unit,out);
		
	out;
	}
	
	
	
```


# EXAMPLE
```{r}
mb.eff = benchmark.efficiency(mb.res);
mb.eff;
```



# trimMe

## My wrapper function

```{r}



trimMe = function(str, side="both", method="stringi")
  {
  # sides = c("both", "left", "right")
  # methods = c("cpp", "stringi", "base")
  side = tolower(side);
  ###m = substr(tolower(method),1,1);  # is this slowing me down?
  # stringr::str_trim(str);
  # if(!is.element(side,sides)) { stop("option for 'side' must be one of:  both, left, right"); }
  # set default to both

  # dump stringr and go to stringi
  # dump tidyverse altogether
  # review httr and build appropriate functions
  # use base whenever possible, then the best packages whenever possible
  # define best by "least associated with new grammar tactics"
  # new grammar is arbitrary and not c-based
  # human-readable functions with variadic inputs is not new grammar
# lsf.str("package:stringi")
# ls("package:stringi")

	if(method == "cpp" && exists("cpp_trim"))
			{
			res = switch(side,
							  "left"  = cpp_ltrim(str),
							  "right" = cpp_rtrim(str),
							  "both"  = cpp_trim(str),
					cpp_trim(str)
					);
			return (res);
			}
		
		
  if( method == "stringi" && isTRUE(requireNamespace("stringi", quietly = TRUE)) )
    {
    res = switch(side,
						  "left"  = stringi::stri_trim_left(str),
						  "right" = stringi::stri_trim_right(str),
						  "both"  = stringi::stri_trim_both(str),
				stringi::stri_trim_both(str)
				);
	return (res);
	}
	
	
		
    res = switch(side,
						  "left"  = gsub("^\\s+", "", str),
						  "right" = gsub("\\s+$", "", str),
						  "both"  = gsub("^\\s+|\\s+$", "", str),
                  gsub("^\\s+|\\s+$", "", str)
                  );
    return (res);
  }


```


## C++
```{r}

# C:\\_git_\\github\\MonteShaffer\\humanVerse\\cpp

Rcpp::sourceCpp("C:\\_git_\\github\\MonteShaffer\\humanVerse\\HVcpp\\src\\str.R.cpp");

### Rcpp::sourceCpp("https://raw.githubusercontent.com/MonteShaffer/humanVerse/main/humanVerse/src/cpp/str.cpp"); # DOESN"T WORK

cpp_trim(" monte \t\n");

```

## MB
```{r}
# http://adv-r.had.co.nz/Rcpp.html
# https://stackoverflow.com/questions/216823/how-to-trim-a-stdstring
# https://teuder.github.io/rcpp4everyone_en/180_date.html#operators-date
# https://gallery.rcpp.org/articles/a-first-boost-example/

str = " jdfldasj djslkfj     ";
# The available units are nanoseconds ("ns"), microseconds ("us"), milliseconds ("ms"), seconds ("s") and evaluations per seconds ("eps") and relative runtime compared to the best median time ("relative").
#.Options$microbenchmark.unit = "eps";
options(microbenchmark.unit="ns");

mb.res = microbenchmark::microbenchmark(
                                        cpp_trim(str), # c++
                                        trimMe(str, method="cpp"),
                                        
                                        
                                        stringi::stri_trim_both(str), 
                                        trimMe(str, method="stringi"),
                                        stringr::str_trim(str, side="both"),
                                        
                                        trimMe(str, method="base"),
                                        
                                        times=10000,
                                        control=list("order"="random","warmup"=100)
                                        );

mb.res = setAttribute("microbenchmark.unit",as.character(options("microbenchmark.unit")),mb.res);

mb.res;
```

### boxplot
```{r}
boxplot(mb.res, horizontal=TRUE);
```

### efficiency
```{r}
mb.eff = benchmark.efficiency(mb.res);
mb.eff;
```

## functional efficiency

```{r}



trimMe = function(str, side="both", method="stringi")
  {
  # sides = c("both", "left", "right")
  # methods = c("cpp", "stringi", "base")
  side = tolower(side);
  ###m = substr(tolower(method),1,1);  # is this slowing me down?
  # stringr::str_trim(str);
  # if(!is.element(side,sides)) { stop("option for 'side' must be one of:  both, left, right"); }
  # set default to both

  # dump stringr and go to stringi
  # dump tidyverse altogether
  # review httr and build appropriate functions
  # use base whenever possible, then the best packages whenever possible
  # define best by "least associated with new grammar tactics"
  # new grammar is arbitrary and not c-based
  # human-readable functions with variadic inputs is not new grammar
# lsf.str("package:stringi")
# ls("package:stringi")

	if(method == "cpp" && exists("cpp_trim"))
			{
			res = switch(side,
							  "left"  = cpp_ltrim(str),
							  "right" = cpp_rtrim(str),
							  "both"  = cpp_trim(str),
					cpp_trim(str)
					);
			return (res);
			}
		
		
  if( method == "stringi" && isTRUE(requireNamespace("stringi", quietly = TRUE)) )
    {
    res = switch(side,
						  "left"  = stringi::stri_trim_left(str),
						  "right" = stringi::stri_trim_right(str),
						  "both"  = stringi::stri_trim_both(str),
				stringi::stri_trim_both(str)
				);
	return (res);
	}
	
	
		
    res = switch(side,
						  "left"  = gsub("^\\s+", "", str),
						  "right" = gsub("\\s+$", "", str),
						  "both"  = gsub("^\\s+|\\s+$", "", str),
                  gsub("^\\s+|\\s+$", "", str)
                  );
    return (res);
  }


```

### Test 1: 'tolower' and 'substr'

```{r}
str = "\t Albjlkje \n\t";
side = "BoTh";
method = "stRINGi";


f1.s = function(side) { tolower(side); }
f1.m = function(method) { tolower(method); }
f2.s = function(side) { substr(tolower(side),1,1); }
f2.m = function(method) { substr(tolower(method),1,1); }

###################################################
options(microbenchmark.unit="ns");

mb.res = microbenchmark::microbenchmark(
                                        f1.s(side),
                                        f1.m(method),
                                        f2.s(side),
                                        f2.m(method),
                                        
                                        times=10000,
                                        control=list("order"="random","warmup"=100)
                                        );

mb.res = setAttribute("microbenchmark.unit",as.character(options("microbenchmark.unit")),mb.res);

mb.res;

###################################################

boxplot(mb.res, horizontal=TRUE);

###################################################

mb.eff = benchmark.efficiency(mb.res);
mb.eff;

```


### Test 2:  'if' (order and nesting)
```{r}
if1 = function(method) { method=="cpp"; }
if2 = function(m) { m=="c"; }
if3t = function(method) { ( method == "cpp" && exists("cpp_trim") ); }
if3tn = function(method) 
          { 
          if( method == "cpp") 
            { 
            if( exists("cpp_trim") ) 
              { 
              return(TRUE); 
              } 
            } 
          return(FALSE); 
          }
if3tr = function(method) { ( exists("cpp_trim") && method=="cpp" ); }
if3trn = function(method) 
          { 
          if( exists("cpp_trim") ) 
            { 
            if( method == "cpp" ) 
              { 
              return(TRUE); 
              } 
            } 
        return(FALSE); 
        }
if3f = function(method) { ( method == "cpp" && exists("cpp_trimDFDS") ); }
if3fr = function(method) { ( exists("cpp_trimDFDS") && method=="cpp" ); }



str = "\t Albjlkje \n\t";
side = "both";
method = "stringi";
s = substr(tolower(side),1,1);
m = substr(tolower(method),1,1);



###################################################
options(microbenchmark.unit="ns");

mb.res = microbenchmark::microbenchmark(
                                        if1(method),
                                        if2(method),
                                        if3t(method),
                                        if3f(method),
                                        if3tr(method),
                                        if3fr(method),
                                        if3trn(method),
                                        
                                        times=10000,
                                        control=list("order"="random","warmup"=100)
                                        );

mb.res = setAttribute("microbenchmark.unit",as.character(options("microbenchmark.unit")),mb.res);

mb.res;

###################################################

boxplot(mb.res, horizontal=TRUE);

###################################################

mb.eff = benchmark.efficiency(mb.res);
mb.eff;

```

### Test 3: 'switch'
```{r}

s1 = function(str,side)
      {
      res = switch(side,
      						  "left"  = gsub("^\\s+", "", str),
      						  "right" = gsub("\\s+$", "", str),
      						  "both"  = gsub("^\\s+|\\s+$", "", str),
                        gsub("^\\s+|\\s+$", "", str)
                        );
      return (res);
      }

s1.d = function(str,side)
      {
      res = switch(side,
      						  "left"  = gsub("^\\s+", "", str),
      						  "right" = gsub("\\s+$", "", str),
                        gsub("^\\s+|\\s+$", "", str)
                        );
      return (res);
      }

s1.if = function(str,side)
    {
    if(side == "left") 
      {
      return( gsub("^\\s+", "", str) );
      }
    if(side == "right") 
      {
      return( gsub("\\s+$", "", str) );
      }
    return( gsub("^\\s+|\\s+$", "", str) );
    }

s2= function(str,s)
      {
      res = switch(s,
      						  "l"  = gsub("^\\s+", "", str),
      						  "r" = gsub("\\s+$", "", str),
      						  "b"  = gsub("^\\s+|\\s+$", "", str),
                        gsub("^\\s+|\\s+$", "", str)
                        );
      return (res);
      }


s2.d= function(str,s)
      {
      res = switch(s,
      						  "l"  = gsub("^\\s+", "", str),
      						  "r" = gsub("\\s+$", "", str),
                        gsub("^\\s+|\\s+$", "", str)
                        );
      return (res);
      }

s2.if = function(str,s)
    {
    if(s == "left") 
      {
      return( gsub("^\\s+", "", str) );
      }
    if(s == "right") 
      {
      return( gsub("\\s+$", "", str) );
      }
    return( gsub("^\\s+|\\s+$", "", str) );
    }

str = "\t Albjlkje \n\t";
side = "both";
method = "stringi";
s = substr(tolower(side),1,1);
m = substr(tolower(method),1,1);



###################################################
options(microbenchmark.unit="ns");

mb.res = microbenchmark::microbenchmark(
                                        s1(str,side),
                                        s1.d(str,side),
                                        s1.if(str,side),
                                        s2(str,s),
                                        s2.d(str,s),
                                        s2.if(str,s),
                                        
                                        times=10000,
                                        control=list("order"="random","warmup"=100)
                                        );

mb.res = setAttribute("microbenchmark.unit",as.character(options("microbenchmark.unit")),mb.res);

mb.res;

###################################################

boxplot(mb.res, horizontal=TRUE);

###################################################

mb.eff = benchmark.efficiency(mb.res);
mb.eff;
```



```{r}
trimMe = function(str, side="both", method="stringi")
  {
  # sides = c("both", "left", "right")
  # methods = c("cpp", "stringi", "base")
  side = tolower(side);
  ###m = substr(tolower(method),1,1);  # is this slowing me down?
  # stringr::str_trim(str);
  # if(!is.element(side,sides)) { stop("option for 'side' must be one of:  both, left, right"); }
  # set default to both

  # dump stringr and go to stringi
  # dump tidyverse altogether
  # review httr and build appropriate functions
  # use base whenever possible, then the best packages whenever possible
  # define best by "least associated with new grammar tactics"
  # new grammar is arbitrary and not c-based
  # human-readable functions with variadic inputs is not new grammar
# lsf.str("package:stringi")
# ls("package:stringi")

	if(method == "cpp" && exists("cpp_trim"))
			{
			res = switch(side,
							  "left"  = cpp_ltrim(str),
							  "right" = cpp_rtrim(str),
							  "both"  = cpp_trim(str),
					cpp_trim(str)
					);
			return (res);
			}
		
		
  if( method == "stringi" && isTRUE(requireNamespace("stringi", quietly = TRUE)) )
    {
    res = switch(side,
						  "left"  = stringi::stri_trim_left(str),
						  "right" = stringi::stri_trim_right(str),
						  "both"  = stringi::stri_trim_both(str),
				stringi::stri_trim_both(str)
				);
	return (res);
	}
	
	
		
    res = switch(side,
						  "left"  = gsub("^\\s+", "", str),
						  "right" = gsub("\\s+$", "", str),
						  "both"  = gsub("^\\s+|\\s+$", "", str),
                  gsub("^\\s+|\\s+$", "", str)
                  );
    return (res);
  }




```

# GCD

## R function
```{r}
gcd.lcm = function(x,y)
	{
	a=x;
	b=y;
	while (b != 0)
        {
		t = b;
		b = a %% b;
		a = t;
        }
	list("gcd"=a, "lcm"=(x*y)/a);
  }

gcd.lcm(20,30);

```


```{r}

# C:\\_git_\\github\\MonteShaffer\\humanVerse\\cpp

# Rcpp::sourceCpp("C:\\_git_\\github\\MonteShaffer\\humanVerse\\cpp\\bits.cpp");

```


## C++ function
```{r}


Rcpp::sourceCpp("C:\\_git_\\github\\MonteShaffer\\humanVerse\\HVcpp\\src\\lcm.R.cpp");
```

```{r}

cpp_gcd(20,30);
cpp_lcm(20,30);

cpp_gcd_lcm(20,30);


x=1444764;
y=8766;

gcd.lcm(x,y);
cpp_gcd_lcm(x,y);
cpp_lcm(x,y);
```



## C++ boost function
```{r}


Rcpp::sourceCpp("C:\\_git_\\github\\MonteShaffer\\humanVerse\\HVcpp\\src\\lcm-boost.R.cpp");


computeGCD(20,30);
computeLCM(20,30);


```

### Test 1: (20,30)

```{r}
x=20;
y=30;


###################################################
options(microbenchmark.unit="ns");

mb.res = microbenchmark::microbenchmark(
                                        cpp_gcd(x,y),
                                        cpp_lcm(x,y),
                                        cpp_gcd_lcm(x,y),
                                        computeGCD(x,y),
                                        computeLCM(x,y),
                                        gcd.lcm(x,y),
                                        
                                        
                                        times=100000,
                                        control=list("order"="random","warmup"=100)
                                        );

mb.res = setAttribute("microbenchmark.unit",as.character(options("microbenchmark.unit")),mb.res);

mb.res;

###################################################

boxplot(mb.res, horizontal=TRUE);

###################################################

mb.eff = benchmark.efficiency(mb.res);
mb.eff;

```



### Test 2: (1444764,8766)

```{r}
# https://cemc.uwaterloo.ca/events/mathcircles/2010-11/Winter/Intermediate_Mar2.pdf
x=1444764;
y=8766;


###################################################
options(microbenchmark.unit="ns");

mb.res = microbenchmark::microbenchmark(
                                        cpp_gcd(x,y),
                                        cpp_lcm(x,y),
                                        cpp_gcd_lcm(x,y),
                                        computeGCD(x,y),
                                        computeLCM(x,y),
                                        gcd.lcm(x,y),
                                        
                                        times=100000,
                                        control=list("order"="random","warmup"=100)
                                        );

mb.res = setAttribute("microbenchmark.unit",as.character(options("microbenchmark.unit")),mb.res);

mb.res;

###################################################

boxplot(mb.res, horizontal=TRUE);

###################################################

mb.eff = benchmark.efficiency(mb.res);
mb.eff;

```



# PRIMES

## R function
```{r}
prime.pracmaPrimes = function(n, first=FALSE)
	{
	# this duplicates the primary logic of pracma::primes
	# by computing 'sqrt' one time, it speeds up things 'slightly'
	# it allows for firstN or fromN with first=FLAG
	gn = n;
	if(first) { gn = ceiling( n * log(n) + n * log(log(n)) ); }
	gn.sqrt = floor( sqrt(gn) );  # needs to round down so the "seq by k" doesn't break ...

	p = seq(1, gn, by = 2); # odd numbers
    q = length(p);
    p[1] = 2; 	# replace 1 with 2 (prime)
				# 9 is the first non-prime?
	if(gn >= 9)
		{
		for (k in seq(3, gn.sqrt, by = 2) )
			{
			k.idx = (k + 1)/2;
			if (p[k.idx] != 0)
				{
				# using a squared rule on indexes ?
				k2.idx = (k * k + 1)/2;

				# cat("\n", " k = ", k, "... (k+1/2) = ", k.idx, "... (k * k + 1)/2 = ", k2.idx, "\n");

				p[ seq(k2.idx, q, by = k) ] = 0;
				}
			}
		}

	p = p[p > 0];

	if(first)
		{
		p[1:n];
		} else {
				p[p < n];
				}
	}

x1 = prime.pracmaPrimes(100);
y1 = prime.pracmaPrimes(100, TRUE);
```



## C++ function
```{r}


Rcpp::sourceCpp("C:\\_git_\\github\\MonteShaffer\\humanVerse\\HVcpp\\src\\primes.R.cpp");



```

```{r}

x2 = cpp_primes(100);
y2 = cpp_primes(100, TRUE);

```


##### TIMER stuff
```{r}

.GlobalEnv$.humanVerse = list();


timer.start = function(key="DEFAULT")
	{
	# this overwrites without any checks?
	.GlobalEnv$.humanVerse[["timer"]][[key]] = list();
	.GlobalEnv$.humanVerse[["timer"]][[key]]$start = Sys.time();  # vs. proc.time()
	}
	
timer.stop = function(key="DEFAULT")
	{
	if(!exists(key, .GlobalEnv$.humanVerse[["timer"]]))
		{
		stop(paste0("Nothing to stop as timer.start for key: ", key, " not called yet!"));
		}
	.GlobalEnv$.humanVerse[["timer"]][[key]]$stop = Sys.time();
	
	.GlobalEnv$.humanVerse[["timer"]][[key]]$diff = as.numeric(.GlobalEnv$.humanVerse[["timer"]][[key]]$stop)-as.numeric(.GlobalEnv$.humanVerse[["timer"]][[key]]$start);
	}


timer.print = function(key="DEFAULT", return="seconds")
	{
	# seconds, pretty-seconds, pretty ...
	if(!exists(key, .GlobalEnv$.humanVerse[["timer"]]))
		{
		stop(paste0("Nothing to print as timer.start/timer.stop for key: ", key, " not called yet!"));
		}
	
	seconds = .GlobalEnv$.humanVerse[["timer"]][[key]]$diff;
		
	if(return == "pretty-seconds") { return( timer.pretty.seconds(seconds) ); }
	
	if(return == "pretty") { return( timer.pretty(seconds) ); }
	
	# DEFAULT
	seconds;
	}
	
	
timer.pretty.seconds = function(seconds)
	{
	paste0( round(seconds,2), " seconds");	
	}

# https://stackoverflow.com/questions/572049/convert-seconds-to-days-minutes-and-hours-in-obj-c
timer.pretty = function(seconds)
	{
				str = "";
	days = floor( seconds / (60 * 60 * 24) );
		seconds = seconds - days * (60 * 60 * 24);
				dstr = "days"; if(days == 1) { dstr = "day"; }
				if(days > 0) { str = paste0(str, days," ",dstr,", "); }
	hours = floor( seconds / (60 * 60) );
				hstr = "hours"; if(hours == 1) { hstr = "hour"; }
				if(days > 0 | hours > 0) { str = paste0(str, hours," ",hstr,", "); }
		seconds = seconds - hours * (60 * 60);
	minutes = floor( seconds / 60 );
				mstr = "minutes"; if(minutes == 1) { mstr = "minute"; }
				if(days > 0 | hours > 0 | minutes > 0) { str = paste0(str, minutes," ",mstr,", "); }
		seconds = seconds - minutes * (60);

	paste0( str, round(seconds,2), " seconds");		
	}









```


### Test 1: (10^3)

```{r}
timer.start("TEST-1");
# > timer.print("TEST-1", return="pretty")
# [1] "2.05 seconds"

n = 10^3;
###################################################
options(microbenchmark.unit="ns");

mb.res = microbenchmark::microbenchmark(
                                        prime.pracmaPrimes(n, FALSE),
                                        prime.pracmaPrimes(n, TRUE),
                                        cpp_primes(n, FALSE),
                                        cpp_primes(n, TRUE),
                                        
                                        times=1000,
                                        control=list("order"="random","warmup"=10)
                                        );

mb.res = setAttribute("microbenchmark.unit",as.character(options("microbenchmark.unit")),mb.res);

mb.res;

timer.stop("TEST-1");
timer.print("TEST-1");
###################################################

boxplot(mb.res, horizontal=TRUE);

###################################################

mb.eff = benchmark.efficiency(mb.res);
mb.eff;


(ceiling( n * log(n) + n * log(log(n)) ));

```


### Test 2: (10^5)

```{r}
timer.start("TEST-2");
# > timer.print("TEST-2", return="pretty") # 10^6 at times = 1000
# [1] "19 minutes, 17.27 seconds"

n = 10^5;
(ceiling( n * log(n) + n * log(log(n)) ));
###################################################
options(microbenchmark.unit="ns");

mb.res = microbenchmark::microbenchmark(
                                        prime.pracmaPrimes(n, FALSE),
                                        prime.pracmaPrimes(n, TRUE),
                                        cpp_primes(n, FALSE),
                                        cpp_primes(n, TRUE),
                                        
                                        times=100,
                                        control=list("order"="random","warmup"=10)
                                        );

mb.res = setAttribute("microbenchmark.unit",as.character(options("microbenchmark.unit")),mb.res);

mb.res;

timer.stop("TEST-2");
timer.print("TEST-2");
###################################################

boxplot(mb.res, horizontal=TRUE);

###################################################

mb.eff = benchmark.efficiency(mb.res);
mb.eff;

```


### LIMITS

```{r}
res = NULL;
N = 10^(1:7);

timer.start("LIMIT");

for(n in N)
  {
  gn = (ceiling( n * log(n) + n * log(log(n)) ));
  
  cat("\n", "n = ",n, "\t", "gn = ",gn,"\n");
  
  timer.start("A");
  A = cpp_primes(n, TRUE);
  timer.stop("A");
  
  timer.start("B");
  B = prime.pracmaPrimes(n, TRUE);
  timer.stop("B");
  
  row = c(n, gn, length(A), max(A), timer.print("A")+1, length(B), max(B), timer.print("B")+1 );
  res = rbind(res, row);
  
  }

res = data.frame(res);
  rownames(res) = N;
  colnames(res) = c("n", "gn", "lengthA", "maxA", "timeA", "lengthB", "maxB", "timeB");
res;

timer.stop("LIMIT");
timer.print("LIMIT", "pretty");

```


#### SCALING
```{r}
x = log10(res$n);
A = res$timeA;
B = res$timeB;
maxT = max(A,B);

plot(x,A, ylim=c(0,maxT), type="b", bty="n", cex=2, lwd=2, pch=1, col="#ff0000", main="", xlab="10^n", ylab="Time (seconds)");
par(new=TRUE);
plot(x,B, ylim=c(0,maxT), type="b", bty="n", cex=2, lwd=2, pch=2, col="#0000ff", main="", xlab="", ylab="");
legend("topleft", c("A: C++","B: R"), pch=c(1,2), col=c("#ff0000","#0000ff"), text.col=c("#ff0000","#0000ff"), cex=2, lwd=2, pt.cex=2, bty="n" );

```
```{r}

par("pty"="s");
plot(x, log10(res$gn), xlim=c(0,9), ylim=c(0,9) );

```



#### EFFICIENCY
```{r}
round( 100* (A-B)/A  , 3);

```







# COMMENT on PRIMES

This bit-wise approach seems best as it minimizes storage space.  The index (location) of the bit is the "prime" value (maybe -1 because of C++ arr[0] indexing).

The lookup in a range would also be faster ... build a hash-map of bits ...




